---
title: "p8105_HW1_jgc2157"
author: "Gunnar"
date: "9/23/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

loading libraries
```{r}
library(tidyverse)
```

*Problem 1*
Creating a data frame with: 
    a random sample of size 10 from a standard Normal distribution
    a logical vector indicating whether elements of the sample are greater than 0
    a character vector of 10 letters
    a factor vector of length 10, with 3 different factor levels - dog, cat, mouse
    
```{r}
problem1_df <- tibble(rsample = rnorm(10),
                      greater_than_zero = rsample > 0,
                      letters_list = letters[1:10],
                      animals = rep(as_factor(c(1, 2, 3)), length.out = 10))
problem1_df
```

Attempting to pull the means of each variable from problem1_df
```{r}
problem1_df %>%
  pull(rsample) %>%
  mean()

problem1_df %>%
  pull(greater_than_zero) %>%
  mean()

problem1_df %>%
  pull(letters_list) %>%
  mean()

problem1_df %>%
  pull(animals) %>%
  mean()
```

Applying as.numeric to the logical, character, and factor variables in problem1_df
```{r}
problem1_df %>%
  pull(greater_than_zero) %>%
  as.numeric()

problem1_df %>%
  pull(letters_list) %>%
  as.numeric()

problem1_df %>%
  pull(animals) %>%
  as.numeric()
```
The logical variable when converted into a numeric variable becomes a binary variable of 0 and 1 because 1 represents TRUE and 0 represents FALSE. This is why a mean can be found for a logical vector.

The character vector when converted into a numeric variable returns a vector of missing values as characters do not correspond to a numeric value. A mean cannot be found for a character vector because of this. 

The factor vector when converted into a numeric variable returns a numeric vector reflecting the levels of the factor. 
