---
title: "p8105_HW1_jgc2157"
author: "Gunnar"
date: "9/23/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

loading tidyverse
```{r}
library(tidyverse)
```

# *Problem 1*
Creating a data frame with: 
    a random sample of size 10 from a standard Normal distribution
    a logical vector indicating whether elements of the sample are greater than 0
    a character vector of 10 letters
    a factor vector of length 10, with 3 different factor levels - dog, cat, mouse
    
```{r}
problem1_df <- tibble(rsample = rnorm(10),
                      greater_than_zero = rsample > 0,
                      letters_list = letters[1:10],
                      animals = rep(as_factor(c("dog", "cat", "mouse")), length.out = 10))
problem1_df
```

Attempting to pull the means of each variable from problem1_df
```{r}
problem1_df %>%
  pull(rsample) %>%
  mean()

problem1_df %>%
  pull(greater_than_zero) %>%
  mean()

problem1_df %>%
  pull(letters_list) %>%
  mean()

problem1_df %>%
  pull(animals) %>%
  mean()
```

Applying as.numeric to the logical, character, and factor variables in problem1_df
```{r, results=FALSE}
problem1_df %>%
  pull(greater_than_zero) %>%
  as.numeric()

problem1_df %>%
  pull(letters_list) %>%
  as.numeric()

problem1_df %>%
  pull(animals) %>%
  as.numeric()
```
The logical variable when converted into a numeric variable becomes a binary variable of 0 and 1 because 1 represents TRUE and 0 represents FALSE. This is why a mean can be found for a logical vector.

The character vector when converted into a numeric variable returns a vector of missing values as characters do not correspond to a numeric value. A mean cannot be found for a character vector because of this. 

The factor vector when converted into a numeric variable returns a numeric vector reflecting the levels of the factor which are character values. As factor levels are not numeric values, a mean cannot be found for them. 

# *Problem 2*

Loading penguin data set
```{r}
data("penguins", package = "palmerpenguins")
```

Summary of penguin data set
```{r}
summary(penguins)

nrow(penguins)
ncol(penguins)
```

*Description of the penguins data set.*
The penguins data set has 8 variables: species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass)g, sex, and year. The data set also has 344 rows. The data set contains information on three bird species (Adelie, Chinstrap, and Gentoo) from three islands (Biscoe, Dream, and Torgersen). The mean flipper length for all birds is 200.9 mm. 

*Scatter plot of flipper length by bill length with color indicating species.*
```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point(size = 1) +
  ggtitle("Flipper Length by Bill Length of Three Bird Species") +
  xlab("Bill Length (mm)") +
  ylab("Flipper Length (mm)")

ggsave("birds_figure1.png")
```

